%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Conclusions.tex                                     %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: Francisco Castro                                           %
%     Last modified :  24 Fev 2026                                      %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Conclusions}
\label{chapter:conclusions}

The work that was developped in this thesis has reasonably met the goals that were stated in section \ref{section:objectives}. Extensive testing of the compliance of the control algorithms to real-time requirements were carried out, which confidently ensures this algorithm is suitable for real-time applications in rough terrain. The mobile robot has rarely been on potentially unsafe poses or scenarios that could compromise either the controller feasibility or the vehicle integrity.

Although the control algorithms were tested under noisy, simulated conditions, this code should be tested on a real-world experiment for the sake of validation. This is one of the main deficiencies of this thesis. 

Moreover, the traversability map generation did not take into account the mobile robot's dynamic features. Apart from that, the fast marching method only computed paths with respect to one goal point. A multi-goal point scenario that resembles several locations of interest within a planetary environment could be envisioned.


% ----------------------------------------------------------------------
\section{Future Work}
\label{section:future}

From this short review of the dissertation main technical deficits, some future research directions are suggested as follows:

\begin{enumerate}
\item This thesis has just presented controllers that include the dynamics model of rigid vehicles. However, rough, uneven, planetary terrains are typically crossed by articulated mobile robots. Therefore, it is suggested that this thesis (or part of it, at least the control sections) is redone under this principle. The Leo-rover presents a simple, reliable and rough-terrain tailored choice to start with \footnote{Leo-rover: https://www.leorover.tech/}. There are also \acrshort{ros} packages made for this rover;
\item Instead of relying on the \texttt{scikit-fmm} Python package to solve the Eikonal equation, one could solely contribute with a custom multi-source solution of the Eikonal equation. \texttt{scikit-fmm} only allows to work with one source, but planetary terrains have multiple locations of interest. If only one source is selected at a time, the global path planner needs to recompute the Eikonal equation \textit{in situ} in order to provide a path to the new goal. If the Eikonal equation can work with several sources at once, only one computation of the potential field is required. In the literature, this approach is called \textit{Fast marching firework method};
\item Gazebo has not proven to be a reliable wheeled robot simulator test-bed at rough terrains. Project Chrono \footnote{Project Chrono: https://projectchrono.org/} (open-source) and Vortex Studio \footnote{Vortex Studio: https://www.cm-labs.com/en/vortex-studio/} (commercial software) are more matured simulators for these terrains. The interaction between rigid wheels and deformable terrain are modeled by them. Reportedly, \acrshort{ros} bridges to these softwares exist. However, care shall be taken as little is known regarding the capability of these softwares to reliably implement and test real-time control algorithms;
\item Real-world experimentation is also an interesting add-on for this kind of work;
\item In this thesis the global path generated by chapter \ref{chapter:globalpathplanner} was not fed to the control algorithm, as the subjects of global path planning and safe real-time control were kept separate and independent of one another. A future work may focus on joining both modules;
\item The virtual speed control module has a lot of room for improvement. In this thesis, the module has only considered the path curvature and distance to the first reference node, but other parameters such as the safe balls radius can be included. Moreover, a variable virtual speed between consecutive reference nodes can be implemented as well;
\item The \textit{online} determination of the terrain parameters is a typical research subject that was not addressed in this thesis. Optimal parameter estimation can be used to estimate them \cite{diehlnumericalocp}. Then, the determined parameters can be fed to the \acrshort{ocp} module, and the vehicle's response to a controller that is unaware of the exact terrain parameters can be studied.
\end{enumerate}


