%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Acknowledgments.tex                                 %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: Francisco Castro                                           %
%     Last modified :  24 Fev 2026                                      %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{\acknowledgments}

% Add entry in the table of contents as section
\addcontentsline{toc}{section}{\acknowledgments}

First of all, I want to thank the Catholic Church and God Himself for all the teachings and continuous spiritual guidance He provides to me since I am born.

Then, I want to thank my beloved parents, brother and grandparents. Throughout all these years, their support, advice and presence can never be properly thanked. If I became who I am, it is solely due to them. Only them know how much it took me to sucessfully complete my studies at IST. 

Secondly, I want to thank my advisor Prof. Rodrigo Ventura. I cannot put into words the availability, sympathy, patience and guidance he provided to me. He has shed light on me when I needed the most and the results of this work are dedicated to him as well.

Then, I want to thank my colleagues and friends Rafael Borges, João Santos, David Santos, Paulo Marques, Diogo Sequeira, José Rodrigues, Pedro Afonso, João Canas, Rui Correia, João Gonçalves, Daniel Schiffart, Joaquim Marques and all the people I worked with at AeroTéc. Their inputs and presence have enriched the time I spent at IST.

I must greatly thank IST for the opportunity to have put myself into such a demanding education system, that has made me into a more matured individual and technically ready to serve as an aerospace engineer. Moreover, the accomodation IST has provided to me as a long-distance student was also of great help during the time I spent in Lisbon.

I also thank all the teachers and professors I had in or outside school. To everyone who has taught me something, I acknowledge your contribution to the individual I am today. To my high school friends Rui, Joel and Tiago as well as my neighbour Adérito Guedes and engineer Nuno Silva, I extend my gratitude for their friendship and for playing a key role on my decision to study aerospace engineering at IST.

